Disk Heating, Galactoseismology, and the Formation of Stellar Halos

### 2017-06-23 paper/paper_v1.tex
copy in kvj's changes

% APW COMMENT: OMG I hate the formatting and math style of this journal, but I
% guess we'll have to just roll with it...

% APW COMMENT: We might get into a battle over the Oxford comma again, but this
% time you're in control!!

% APW COMMENT: Stylistic comment: I don't like sentences like "\cite{...} found
% that XX" or "As found in \cite{...}" -- my personal rule is that I should be
% able to read the text if I deleted all of the \cite commands. So, I instead
% prefer to write summary statements, and then use \citep{...} as the reference
% at the end of a given statement. I tried to change instances I could found,
% but that might be controversial and I hope I don't annoy you too much...

% APW COMMENT: We should mention Hercules-Aquila as a genuine "shell" candidate
% since it does have RR Lyrae?

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsrefs}
% \usepackage{aas_macros} % APW: I needed this for the references...

@@

% Authors, for the paper (add full first names)
\Author{Kathryn Johnston $^{1,\dagger,\ddagger}$, Firstname Lastname $^{1,\ddagger}$ and Sanjib Sharma $^{2,}$*}
% Rachel Beaton? Steve Majewski? Sanjib Sharma?

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: +x-xxx-xxx-xxxx}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3}
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

@@

Emerging in the 1990's, the catalogues that inspired this recent acceleration not only mapped global structures in our Galaxy, but also revealed the ubiquity of substructure within it.
These revelations added an unforeseen richness to interpretations, and encouraged the development of new dynamical tools for studying ongoing interactions and formation histories:
\survey{Hipparcos} \citep{esa97} found moving groups of stars in the velocity distribution in the Solar Neighborhood \citep{dehnen98}, some corresponding (as expected) to destroyed stars clusters, while others (unexpectedly) could plausibly be explained as signatures of resonances with the Galactic bar \citep{dehnen00};
the Sloan Digital Sky Survey \citep[hereafter, SDSS ---][]{york00,stoughton02,abazajian03} found many streams of Main Sequence Turnoff (MSTO) stars in our stellar halo, thought to be from long-dead satellite galaxies and dissolved globular clusters \citep{newberg02,belokurov06} as a stunning confirmation that our Galaxy had indeed formed hierarchically \citep[e.g.][]{bullock01,bullock05};
and M giant stars selected from the Two Micron All Sky Survey \citep[hereafter, 2MASS ---][]{nikolaev00} traced the debris from the Sagittarius dwarf galaxy entirely around the Sky \citep{majewski03} to offer a window on the history of its disruption \citep{law10}.
From these and many other studies that use large survey catalogues, it is now clear that the Milky Way is full of kinematic substructure.

% APW COMMENT: this section makes it sound like the surveys discovered these
% things "for free," but really it was the authors you cite who found these
% things. I would rearrange to say, e.g., "Using data from the
% \survey{Hipparcos} mission \citep{esa97}, moving groups of stars were found
% in the velocity distribution in the Solar Neighborhood
% \citep{dehnen98}...etc"

\begin{figure}[t]
\label{fig:ting}
\centering
%\includegraphics[width=3 cm]{logo-mdpi}
\caption{\label{fig:ting}
Summary of spatial distributions of M giants
The distance measurements are from photometry alone and have expected absolute uncertainties of $\approx$20\% \citep{sheffield14}.
}
\end{figure}
% APW COMMENT: Maybe Ting would be willing to share her data, and I can re-make
% the figure? You know how I am about visualizations... :)

@@

%Any fine features discussed below will not be apparent.
Each of these substructures were originally identified using SDSS and 2MASS photometry as over-densities in stellar number counts relative to the expected global structure of the disk or inner stellar halo.
Unlike stellar streams, these substructures are present at a range of low to moderate Galactic latitudes --- we will hereafter refer to them collectively as the ``low-latitude structures.''
Note that the same region on the sky has been shown to be richly structured on even smaller scales\citep{slater14,martin14,deason14} but we do not discuss these sub-substructures here.

@@

\item{\it The Monoceros Ring or Giant Anticenter Stellar Structure} (hereafter Mon/GASS) is an arc- or ring-like feature of stars beyond the expected edge of the Galactic disk \citep[assumed to be $\approx 5~\kpc$ beyond the Sun in cylindrical radius][]{robin92}.
The feature spans a large area: Galactic longitudes $\approx 120^\circ \lesssim l \lesssim 240^\circ$, latitudes $-30^\circ \lesssim b \lesssim +40^\circ$, and Heliocentric distances $5\lesssim d_\odot \lesssim 10~\kpc$ (\citealt{Morganson:2016}).
Mon/GASS was originally found as an excess of main sequence stars in SDSS \citep{yanny03,ibata03} and M giant stars in 2MASS \citep{rochapinto03}.
Follow-up spectra of the M giant stars indicate that they follow a clear trend in mean velocity with Galactic longitude with small dispersion \citep{crane04}.
\item{\it The Triangulum-Andromeda Clouds I and II} (hereafter TriAnd 1 and 2) were first discovered as a single diffuse over-density of M giant stars covering the area $\approx 100^\circ \lesssim l \lesssim 150^\circ$, $\approx 20^\circ \lesssim b \lesssim 40^\circ$ overlapping the Mon/GASS structure on the sky but at larger Heliocentric distances of approximately $15$--$30~\kpc$ \citep{rochapinto04}.
Follow-up spectroscopy showed that the M giants exhibit a coherent velocity sequence with a small dispersion \citep{rochapinto04} and later deep photometry in the region revealed associated MSTO stars and an estimate of the surface brightness of the feature $\Sigma > 32~{\rm mag}~{\rm arcsec}^{-2}$ \cite{majewski04}.
Subsequent photometric follow-up then revealed the presence of {\it two} distinct main sequences \citep{martin07}.
\item{\it A13} is another tenuous association of stars discovered by applying a group finding algorithm \citep{sharma09} to the catalogue of all 2MASS M giants \citep{sharma10}.

@@

Mon/GASS has been attributed to the accretion of a satellite \citep[on a retrograde orbit][]{penarrubia05, michel-dansac11}, a natural extension of the Galactic warp \citep{momany04,momany06}, or disturbances to the Galactic disk \citep{kazantzidis08,younger08,purcell11,xu15,gomez16}.
For TriAnd I/II and A13, their extreme positions at $(R,Z) \approx (20, -8)~\kpc$, $(R,Z) \approx (25, -10)~\kpc$, and $(R,Z) \approx (20, 8)~\kpc$ (across a large range in Galactocentric azimuth, $\phi$) seems to exclude a disturbed disk as a possible origin in these cases, and a satellite on a retrograde orbit has again been shown to provide a plausible explanation \citep{sheffield14}.

@@

A more convincing, coherent picture of the nature of these low-latitude structures is just beginning to emerge.
This paper reviews the contributions our own group has made to formulating this picture, which include spectroscopic surveys of the low-latitude structures to study metallicity and velocities \citep{sheffield14,ting17}, stellar populations \citep{pricewhelan15,sheffield17}, and abundance patterns \citep{bergemann17}, as well as numerical simulations \citep{sheffield14,laporte17a,laporte17b}.
We summarize this observational and theoretical work in Section 2 and 3 respectively, adding in the context of contemporary work from other groups, as well as the larger context of possible connections across the Galactic disk.
Armed with this understanding of the nature of these substructures, we go on in Section 4 to discuss prospects for mapping such structures more generally around our own and other Galaxies.
We end in Section 5 by outlining the motivation for making such maps, asking what they might be telling us about bigger questions: galaxy formation scenarios and dark matter distributions.

% APW COMMENT: I'm not sure about the subsection grouping below.  I'm wondering
% if there is another way to reorganize...Perhaps subsections for name of
% structure, subsubsections for "observational follow-up", "theoretical
% understanding", "future work needed"? i.e. take the transpose of what is here
% now, if you know what I mean.

@@

From clustering in positions or distance alone, many candidate groups and
over-densities of M giants have been identified in the outer disk / inner halo.
Over the last five years our group has pursued a spectroscopic campaign to
follow up members of these candidate structure with the aims of (1) confirming
the existence of substructure in velocities, (2) measuring chemical abundances,
and (3) studying the constituent stellar populations.
These goals then inform our own efforts to produce plausible dynamical
formation scenarios using simulations.
We were interested in particular in {\it avoiding} the collimated stellar streams that have been well-studied in prior work \citep[such as Sgr, Orphan, GD1 and Pal 5 --- see, e.g.,][]{law10,koposov10,kuepper15,bovy16}
and instead considered groups that appeared as diffuse, amorphous, extended stellar structures such as TriAnd 1 and 2 \citep{rochapinto04} and A13 \citep{sharma10}.
Initial interpretations of these morphologies suggested the structures could be {\it shells} --- debris from the disruption of satellite galaxies on near-radial orbits \citep{johnston08} --- as seen from an internal perspective.

In our first results \citep{sheffield14}, we extended a prior sample of TriAnd
M giants \citep{rochapinto04} by obtaining spectra of all members identified
with an automated clustering algorithm applied to 2MASS catalog sources in the
region \citep{sharma10}.
We found that the groups members formed two sequences in color-magnitude space, which we identified with the two distinct MSTO in optical data by \cite{martin07}, named TriAnd 1 and TriAnd 2 respectively.
Both sequences form a clear over-density in radial velocity over the background halo velocity distribution and form a low velocity-dispersion ($\sim 25~\kms$) sequence with a steady gradient of mean Galactic standard-of-rest (GSR) radial velocity ($v_{\rm GSR}$) from positive to negative $v_{\rm GSR}$ with increasing $l$.
% APW COMMENT: I added some more specifics, but it feels a little clunky.
% Trying to say that it goes from positive vGSR at low longitudes to negative
% vGSR at larger longitudes
In this work, we presented a model that simultaneously and approximately reproduces TriAnd 1 and 2 as tidal debris stripped over two separate pericentric passages from a single accreted satellite on a low-eccentricity, retrograde, near-planar orbit.
These debris structures were morphologically closer to {\it streams} rather
than {\it shells}, but still subtended large areas on the sky as observed from
the Sun's position.

\cite{li17} continued the spectroscopic survey to look at A13, again finding coherent velocity structures with low dispersion.
% APW COMMENT: (I'm sure you know) but this paragraph needs more detail about
% the Li et al. paper. Adding this as a TODO for us both.

@@

Figure~\ref{fig:ting_vel} (reproduced from \citealt{li17}), summarizes the line-of-sight (GSR) velocity trends of M giants in each of the three low-latitude structures.
Not only do these groups all have low dispersions relative to the stellar halo
--- which confirms the association of stars within each group --- they also
appear to collectively exhibit a continuous gradient with Galactic longitude,
$l$.
This latter finding suggests that the groups may be associated.

@@

\subsection{Stellar Populations}

Motivated by the observed, low-dispersion velocity distribution of the M giant stars in TriAnd, we sought to observe other distance tracer stars --- RR Lyrae-type pulsators --- in the region, determine their membership, and improve the distance estimates to the structure.
We selected RR Lyrae stars in the TriAnd region from the pre-classified
variable star catalog produced by the Catalina Sky Survey \citep{drake14},
using a conservative distance cut to account for uncertainties in the RR Lyrae
and M giant photometric distance estimates.
We obtained spectra for $\approx$1/3 of the total number of RR Lyrae in the M
giant volume considered to be associated with TriAnd 1 and measured radial
velocities for these stars \cite{pricewhelan15}.

@@

Figure~\ref{fig:apw} \citep[reproduced from][]{pricewhelan15} shows the results of our survey: unlike the M giants (gray points) the RR Lyrae (black points) show no clear, tight velocity sequence.
By modeling both the RR Lyrae and M giants velocity data as mixtures of two gaussians representing a cold foreground sequence and a possible background halo population of larger dispersion, we were able to show that the number ratio of RR Lyrae to M giants (\frrmg) within the overdensity was $<0.38$ with 95\% confidence.

In order to aid with the interpretation of this discovery, Figure~\ref{fig:isochrones} plots isochrones for two $10~{\rm Gyr}$-old populations of low and intermediate metallicity, including the color range for RR Lyrae and M spectral class.
It emphasizes our understanding that these two types of stars are tracers of populations with quite distinct metallicities and illustrates why, while the initial aim of our survey to find more accurate distances to TriAnd I and II had failed, our results could be even more intriguing.
The metallicity distributions in nearly all existing satellite galaxies
orbiting the Milky Way \citep[e.g.,][]{kirby11} are typically biased towards
low metallicities, such that they contain no or few M giant stars (i.e. $\frrmg = \infty$).
The largest satellites (the Large Magellanic Cloud and the Sagittarius dwarf galaxy) are exceptions as they contain the most metal rich populations; these dwarfs have $\frrmg \sim 0.5$ \citep{pricewhelan15}.
In contrast, the Galactic disk is an overall metal-rich population and thus has very few RR Lyrae \citep[i.e. $f_{\rm RR:M} \sim 0$][]{amrose01}, more consistent with our findings for the stellar population of TriAnd 1.

Our work on the stellar populations of the TriAnd region motivated us to look at possible associations of RR Lyrae with the M giant sequences found in Mon/GASS and A13 \citep{sheffield17}.
For the surveyed regions of these structures, we find \frrmg\ values similar to those observed for TriAnd , and consistent with membership of the Galactic disk.

@@

% APW COMMENT: This section needs more context and perhaps much less detail
% about the actual observations...

\cite{chou2010b} took high-resolution spectra of 21 M giants in the GASS overdensity. They found that for some stars the abundances of titanium and yttrium deviate from that of the Milky Way disk stars at the same metallicity.  For Ti, the measured abundances are lower by up to 0.4 dex compared to the mean trends known for main-sequence stars of the Galactic disk (e.g Reddy et al. 2003, Bensby et al. 2014). The mean offset for [Y/Fe] is about 0.2 dex at [Fe/H] $\approx -0.5$. Comparing with their own results in Sgr dwarf spheroidal galaxy \citep{chou2010a} concluded that the abundances are consistent with the GASS stars having an external origin. It should be noted, however, that their results do not agree with the abundances of Ti and, in particular, La, measured in the Canis Major overdensity by \cite{sbordone2005}. Their analysis was based on spectra covering a limited spectral region in the near-IR, including 11 neutral iron and 2 neutral titanium lines. Also \cite{chou2010b} caution that the effects of non-local thermodynamic equilibrium (non-LTE) may be significant, since their measurement of metallicity and Ti abundances relies on the LTE analysis of lines of neutral species,

We have recently obtained high-resolution and signal-to-noise spectra of 15 stars in TriAnd1 and A13 overdensities. 14 stars were observed with the HIRES-S spectrograph at the Keck-1 telescope \cite{vogt1994} and 1 star was observed using the the UVES spectrograph at the VLT (Program ID: 097.B-0770A). The spectral resolving power R of the HIRES spectra is 36\,000 and the UVES data have R $\sim$ 47\,000. All Keck spectra cover the full optical region, from $4800$ to $8770$ $\AA$, although the exact wavelength coverage varies as several slightly different instrument configurations were tried in an effort to get all the key lines into a single exposure. The signal-to-noise (SNR)/\AA of the HIRES spectra near 5200 \AA in the continuum at the center of the echelle order exceeds 200. For the UVES spectrum, the SNR of 50 near 5500 \AA was achieved. The fundamental parameters and chemical abundances of stars were determined using 2MASS and APASS photometry and the high-resolution spectra. All 15 stars are M-type giants with effective temperatures Teff ~ 3800 K and surface gravities log(g) ~ 1 dex. The $T_{\rm eff}$ estimates were derived using the Infrared Flux Method (Casagrande et al. 2010, 2014). Chemical abundances were measured for 6 chemical elements, including O, Na, Mg, Ca, Fe, Eu using LTE and the standard MARCS stellar model atmospheres (Gustafsson et al. 2008). We have also performed detailed analysis of the effects of non-local thermodynamic equilibrium (non-LTE) for the elements, where detailed calculations are available \cite{bergemann2011, bergemann2012, bergemann2016}, however, the NLTE corrections for the chosen lines are minor and do not change our conclusions. In what follows, we use our LTE results, because all studies of chemical abundances in dSph systems and most studies in the Milky Way to date have employed LTE with 1D hydrostatic models

>>>>

% APW COMMENT: OMG I hate the formatting and math style of this journal, but I
% guess we'll have to just roll with it...
% KVJ: agreed
% APW COMMENT: We might get into a battle over the Oxford comma again, but this
% time you're in control!!
% KVJ: OK

% APW COMMENT: Stylistic comment: I don't like sentences like "\cite{...} found
% that XX" or "As found in \cite{...}" -- my personal rule is that I should be
% able to read the text if I deleted all of the \cite commands. So, I instead
% prefer to write summary statements, and then use \cite{...} as the reference
% at the end of a given statement. I tried to change instances I could found,
% but that might be controversial and I hope I don't annoy you too much...
% KVJ: sounds fine to me

% APW COMMENT: We should mention Hercules-Aquila as a genuine "shell" candidate
% since it does have RR Lyrae?
% KVJ: Sure, but I'll leave you to figure out where and how....

\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsrefs}
%\usepackage{aas_macros} % APW: I needed this for the references...

@@

% Authors, for the paper (add full first names)
\Author{Kathryn V Johnston $^{1,\dagger}$, Adrian M Price-Whelan$^{2,\dagger}$, Maria Bergemann$^{3}$, Chervin Laporte$^{1}$, Ting S Li$^{4}$, Allyson A Sheffield$^{5}$, Branimir Sesar$^{3}$ and Sanjib Sharma $^{6,}$}
% Rachel Beaton? Steve Majewski? Sanjib Sharma?

@@

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{
$^{1}$ \quad Department of Astronomy, Columbia University \\
$^{2}$ \quad Department of Astrophysical Sciences, Princeton University \\
$^{3}$ \quad Max Planck Institute for Astronomy, Heidelberg\\
$^{4}$ \quad Department of Physics \& Astronomy, Texas A \& M University\\
$^{5}$ \quad City University of New York, LaGuardia Community College \\
$^{6}$ \quad Sydney Institute for Astronomy, School of Physics, University of Sydney \\
}

% Contact information of the corresponding author
\corres{Correspondence: kvj@astro.columbia.edu; Tel.: +1-212-854-3884}

% Current address and/or shared authorship
%\firstnote{Current address: Affiliation 3}
\firstnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

@@

Emerging in the 1990's, the catalogues that inspired this recent acceleration not only mapped global structures in our Galaxy, but also revealed the ubiquity of substructure within it.
These revelations added an unforeseen richness to interpretations, and encouraged the development of new dynamical tools for studying ongoing interactions and formation histories:
\survey{Hipparcos} \cite{esa97} found moving groups of stars in the velocity distribution in the Solar Neighborhood \cite{dehnen98}, some corresponding (as expected) to destroyed stars clusters, while others (unexpectedly) could plausibly be explained as signatures of resonances with the Galactic bar \cite{dehnen00};
the Sloan Digital Sky Survey \cite[hereafter, SDSS ---][]{york00,stoughton02,abazajian03} found many streams of Main Sequence Turnoff (MSTO) stars in our stellar halo, thought to be from long-dead satellite galaxies and dissolved globular clusters \cite{newberg02,belokurov06} as a stunning confirmation that our Galaxy had indeed formed hierarchically \cite[e.g.][]{bullock01,bullock05};
and M giant stars selected from the Two Micron All Sky Survey \cite[hereafter, 2MASS ---][]{nikolaev00} traced the debris from the Sagittarius dwarf galaxy entirely around the Sky \cite{majewski03} to offer a window on the history of its disruption \cite{law10}.
From these and many other studies that use large survey catalogues, it is now clear that the Milky Way is full of kinematic substructure.

% APW COMMENT: this section makes it sound like the surveys discovered these
% things "for free," but really it was the authors you cite who found these
% things. I would rearrange to say, e.g., "Using data from the
% \survey{Hipparcos} mission \cite{esa97}, moving groups of stars were found
% in the velocity distribution in the Solar Neighborhood
% \cite{dehnen98}...etc"
% KVJ: will get to if there is time

\begin{figure}[t]
\label{fig:ting}
\centering
%\includegraphics[width=3 cm]{logo-mdpi}
\caption{\label{fig:ting}
Summary of spatial distributions of M giants
The distance measurements are from photometry alone and have expected absolute uncertainties of $\approx$20\% \cite{sheffield14}.
}
\end{figure}
% APW COMMENT: Maybe Ting would be willing to share her data, and I can re-make
% the figure? You know how I am about visualizations... :)
% KVJ: that makes me very happy

@@

Each of these substructures were originally identified using SDSS and 2MASS photometry as over-densities in stellar number counts relative to the expected global structure of the disk or inner stellar halo.\footnote{Note that the same region on the sky has been shown to be richly structured on even smaller scales\cite{slater14,martin14,deason14} but we do not discuss these sub-substructures here.}
Unlike stellar streams, these substructures are present at a range of low to moderate Galactic latitudes --- we will hereafter refer to them collectively as the ``low-latitude structures''.
The following is a brief  summary of the discovery and initial follow-up papers that outlined their basic properties.

@@

\item{\it The Monoceros Ring or Giant Anticenter Stellar Structure} (hereafter Mon/GASS) is an arc- or ring-like feature of stars beyond the expected edge of the Galactic disk \cite[assumed to be $\approx 5~\kpc$ beyond the Sun in cylindrical radius ---][]{robin92}.
The feature spans a large area: Galactic longitudes $\approx 120^\circ \lesssim l \lesssim 240^\circ$, latitudes $-30^\circ \lesssim b \lesssim +40^\circ$, and Heliocentric distances $5\lesssim d_\odot \lesssim 10~\kpc$ (\cite{Morganson:2016}).
Mon/GASS was originally found as an excess of main sequence stars in SDSS \cite{yanny03,ibata03} and M giant stars in 2MASS \cite{rochapinto03}.
Follow-up spectra of the M giant stars indicate that they follow a clear trend in mean velocity with Galactic longitude with small dispersion \cite{crane03}.
\item{\it The Triangulum-Andromeda Clouds} (hereafter TriAnd 1 and TriAnd 2) were first discovered as a single diffuse over-density of M giant stars covering the area $\approx 100^\circ \lesssim l \lesssim 150^\circ$, $\approx 20^\circ \lesssim b \lesssim 40^\circ$ overlapping the Mon/GASS structure on the sky but at larger Heliocentric distances of approximately $15$--$30~\kpc$ \cite{rochapinto04}.
Follow-up spectroscopy showed that the M giants exhibit a coherent velocity sequence with a small dispersion \cite{rochapinto04} and later deep photometry in the region revealed associated MSTO stars and an estimate of the surface brightness of the feature $\Sigma > 32~{\rm mag}~{\rm arcsec}^{-2}$ \cite{majewski04}.
Subsequent photometric work also revealed the presence of {\it two} distinct main sequences \cite{martin07}.
\item{\it A13} is another tenuous association of stars discovered by applying a group finding algorithm \cite{sharma09} to the catalogue of all 2MASS M giants \cite{sharma10}.

@@

Mon/GASS has been attributed to the accretion of a satellite on a retrograde orbit \cite{penarrubia05}, a natural extension of the Galactic warp \cite{momany04,momany06}, or disturbances to the Galactic disk \cite{kazantzidis08,younger08,purcell11,xu15,gomez16}.
For the TriAnd clouds, their extreme positions at $(R,Z) \approx (20, -8)~\kpc$ and $(R,Z) \approx (25, -10)~\kpc$, 
%and $(R,Z) \approx (20, 8)~\kpc$ 
(across a large range in Galactocentric azimuth, $\phi$) seems to exclude a disturbed disk as a possible origin in these cases and a satellite on a retrograde orbit has again been shown to provide a plausible explanation of their structure \cite{sheffield14}.
Indeed, preliminary abundance studies show [$\alpha$/Fe] values for member M giants of both Mon/GASS and the TriAnd Clouds reminiscent of stars in Galactic satellites and unlike the disk \cite{chou2010b,chou11} --- see Section \ref{sec:abundances} for a more complete discussion.

@@

% KVJ: hope sentence above helps?

A more convincing, coherent picture of the nature of these low-latitude structures is just beginning to emerge.
This paper reviews the contributions our own group has made to formulating this picture, which include spectroscopic surveys of the low-latitude structures to study metallicity and velocities \cite{sheffield14,li17}, stellar populations \cite{pricewhelan15}, and abundance patterns (paper currently in preparation, Bergemann et al 2017), as well as numerical simulations \cite{sheffield14,laporte16}.
We summarize this observational and theoretical work in Section 2 and 3 respectively, adding in the context of contemporary work from other groups, as well as the larger context of possible connections across the Galactic disk.
Armed with this understanding of the nature of these substructures, we go on in Section 4 to discuss prospects for mapping such structures more generally around our own and other Galaxies.
We end in Section 5 by outlining the motivation for making such maps, asking what they might be telling us about bigger questions: galaxy formation scenarios and dark matter distributions.

% APW COMMENT: I'm not sure about the subsection grouping below.  I'm wondering
% if there is another way to reorganize...Perhaps subsections for name of
% structure, subsubsections for "observational follow-up", "theoretical
% understanding", "future work needed"? i.e. take the transpose of what is here
% now, if you know what I mean.
% KVJ: not _sure_ I agree and no time to implement

@@

From clustering in positions or distance alone, many candidate groups and
over-densities of M giants have been identified in the outer disk / inner halo.
Over the last five years our group has pursued a spectroscopic campaign to
follow up members of these candidate structure with the aims of (1) confirming
the existence of substructure in velocities, (2) measuring chemical abundances,
and (3) studying the constituent stellar populations.
These goals then inform our own efforts to produce plausible dynamical
formation scenarios using simulations.
We were interested in particular in {\it avoiding} the collimated stellar streams that have been well-studied in prior work \cite[such as Sgr, Orphan, GD1 and Pal 5 --- see, e.g.,][]{law10,koposov10,kuepper15,bovy16}
and instead considered groups that appeared as diffuse, amorphous, extended stellar structures such as TriAnd 1 and 2 \cite{rochapinto04} and A13 \cite{sharma10}.
Initial interpretations of these morphologies suggested the structures could be {\it shells} --- debris from the disruption of satellite galaxies on near-radial orbits \cite{johnston08} --- as seen from an internal perspective.

In our first results \cite{sheffield14}, we extended a prior sample of TriAnd
M giants \cite{rochapinto04} by obtaining spectra of all members identified
with an automated clustering algorithm applied to 2MASS catalog sources in the
region \cite{sharma10}.
We found that group members occupied two distinct loci in color-magnitude space, which we associated with the two MSTO identified in optical data by \cite{martin07}, named TriAnd 1 and TriAnd 2 respectively.
Stars in both TriAnd 1 and TriAnd 2  form  clear over-densities in radial velocity with low dispersion ($\sim 25~\kms$) compared to the background halo velocity distribution.
They both follow the same  sequence in velocity with a steady gradient of mean Galactic standard-of-rest (GSR) radial velocity ($v_{\rm GSR}$) from positive to negative $v_{\rm GSR}$ with increasing $l$.
% APW COMMENT: I added some more specifics, but it feels a little clunky.
% Trying to say that it goes from positive vGSR at low longitudes to negative
% vGSR at larger longitudes
%KVJ: slight re-write
In this work, we presented a model that simultaneously and approximately reproduces TriAnd 1 and 2 as tidal debris stripped over two separate pericentric passages from a single accreted satellite on a low-eccentricity, retrograde, near-planar orbit.
The debris structures in the simulations were morphologically closer to {\it streams} rather
than {\it shells}, but still subtended large areas on the sky as observed from
the Sun's position.

In subsequent work we continued the spectroscopic survey to look at A13, the new group M giant discovered through the group-finding algorithm \cite{sharma10}. A13 overlaps the TriAnd clouds in Galactic longitude, but is apparent in the Northern Galactic hemisphere at slightly brighter magnitudes. The spectra showed that, like the TriAnd clouds, this  group had a coherent velocity structure with low dispersion and a steady gradient with $l$, confirming the genuine association of its members \cite{li17}.
% APW COMMENT: (I'm sure you know) but this paragraph needs more detail about
% the Li et al. paper. Adding this as a TODO for us both.
% KVJ: I just added a couple of sentences. Don't want seciton too long
...

@@

Figure~\ref{fig:ting_vel} (reproduced from \cite{li17}), summarizes the line-of-sight (GSR) velocity trends of M giants in each of the three low-latitude structures.
Not only do these groups all have low dispersions relative to the stellar halo
--- which confirms  groups themselves as real --- they also
appear to {\it collectively} exhibit a continuous gradient with Galactic longitude,
$l$.
This latter finding suggests that the groups may also be associated with each other.

@@

\subsection{Stellar Populations}
 \label{sec:populations} 
Motivated by the observed, low-dispersion velocity distribution of the M giant stars in TriAnd, we sought to observe other distance tracer stars --- RR Lyrae-type pulsators --- in the region, determine their membership, and improve the distance estimates to the structure.
We selected RR Lyrae stars in the TriAnd region from the pre-classified
variable star catalog produced by the Catalina Sky Survey \cite{drake14},
using a conservative distance cut to account for uncertainties in the RR Lyrae
and M giant photometric distance estimates.
We obtained spectra for $\approx$1/3 of the total number of RR Lyrae in the M
giant volume considered to be associated with TriAnd 1 and measured radial
velocities for these stars \cite{pricewhelan15}.

@@

Figure~\ref{fig:apw} \cite[reproduced from][]{pricewhelan15} shows the results of our survey: unlike the M giants (gray points) the RR Lyrae (black points) show no clear, tight velocity sequence.
By modeling both the RR Lyrae and M giants velocity data as mixtures of two gaussians representing a cold foreground sequence and a possible background halo population of larger dispersion, we were able to show that the number ratio of RR Lyrae to M giants (\frrmg) within the overdensity was $<0.38$ with 95\% confidence.

%\begin{figure}[t]
%\centering
%\includegraphics[width=3 cm]{logo-mdpi}
%\caption{\label{fig:isochrones}
%Isochrones for various ages metallicities}
%\end{figure}

Since we were unable to find any RR Lyrae clearly associated with TriAnd 1, our attempt to measure a more accurate distance to this structure was a failure.
However, the upper limit on the value of $\frrmg$ was in itself interesting.
RR Lyrae and M giants are tracers of populations with quite distinct metallicities: stars in the horizontal branch phase of evolution are typically only blue enough to cross the instability strip and become RR Lyrae if they have [Fe/H]$\lesssim -1.5$; and giant stars typically only evolve to colors red enough to become spectral class M if they have [Fe/H]$\gtrsim$-1.5.
Hence, a stellar population has to contain a significant range of metallicities in order to contain both types of stars.
%In order to aid with the interpretation of this discovery, Figure~\ref{fig:isochrones} plots isochrones for two $10~{\rm Gyr}$-old populations of low and intermediate metallicity, including the color range for RR Lyrae and M spectral class.
%It emphasizes our understanding that these two types of stars are tracers of populations with quite distinct metallicities and illustrates why, while the initial aim of our survey to find more accurate distances to TriAnd 1 and 2 had failed, our results could be even more intriguing.
The metallicity distributions in nearly all existing satellite galaxies
orbiting the Milky Way \cite[e.g.,][]{kirby11} are typically biased towards
low metallicities, such that they contain no or few M giant stars (i.e. $\frrmg = \infty$).
The largest satellites (the Large Magellanic Cloud and the Sagittarius dwarf galaxy) are exceptions as they contain the most metal rich populations; these dwarfs have $\frrmg \sim 0.5$ \cite{pricewhelan15}.
In contrast, the Galactic disk is an overall metal-rich population and thus has very few RR Lyrae \cite[i.e. $f_{\rm RR:M} \sim 0$][]{amrose01}, more consistent with our findings for the stellar population of TriAnd 1.

Our work on the stellar populations of the TriAnd region motivated us to look at possible associations of RR Lyrae with the M giant sequences found in Mon/GASS and A13.
For the surveyed regions of these structures, we find \frrmg\ values similar to those observed for TriAnd , and consistent with membership of the Galactic disk. A paper on our results is now in preparation (Sheffield et al 2017).

@@

\label{sec:abundances}
% APW COMMENT: This section needs more context and perhaps much less detail
% about the actual observations...

The origin of stellar associations can also be explored through the chemical compositions of their constituent stars.
However, our finding (described in Section \ref{sec:populations}) that the low-latitude-structures all have stellar populations (as indicated by $f_{\rm RR:M}$) that look more like the disk than known Galactic satellites is at odds with prior work on abundance patterns.
For example, high-resolution spectra of 21 M giants in Mon/GASS \cite{chou2010b} showed 
[Ti/Fe] lower by up to 0.4 dex compared to the mean trends known for main-sequence stars of the Galactic disk (e.g Reddy et al. 2003, Bensby et al. 2014) and a mean offset for [Y/Fe] of about 0.2 dex at [Fe/H] $\approx -0.5$. A comparison with results from the same group for stars in the Sgr dwarf spheroidal galaxy \cite{chou2010a} suggested more similar patterns to those seen in Mon/GASS and support  an external origin for this structure. Subsequent work comes to similar conclusions for the TriAnd Clouds \cite{chou11}.
%It should be noted, however, that their results do not agree with the abundances of Ti and, in particular, La, measured in the Canis Major overdensity by \cite{sbordone2005}. Their analysis was based on spectra covering a limited spectral region in the near-IR, including 11 neutral iron and 2 neutral titanium lines. Also \cite{chou2010b} caution that the effects of non-local thermodynamic equilibrium (non-LTE) may be significant, since their measurement of metallicity and Ti abundances relies on the LTE analysis of lines of neutral species,
